Efficient UQ for chemical systems using active \author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan} subspaces

### 2018-09-25 Paper/method.tex
..

As mentioned earlier, the model output $f(\bm{\xi})$ i.e. the igntion delay in the
H$_2$/O$_2$ reaction in this case, varies
 predominantly in the active subspace. Hence, 
$f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
the subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
regarded as the \textit{sufficient summary plot} (SSP), obtained using the perturbation-based and
regression-based gradient estimates are
compared in Figure~\ref{fig:comp_ssp}.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp}
\caption{An illustrative comparison of the SSPs generated using the 
perturbation-based and the regression-based strategies for computing the active subspace.}
\label{fig:comp_ssp}
\end{center}
\end{figure}
%
The two SSPs are observed to be in excellent agreement with each other. Moreover, it is interesting to note
that the response in ignition
delay based on the considered probability distributions for $A_i$'s is observed to be approximately linear in the
active subspace. This observation further indicates that the active subspace is 1-dimensional in this case.

We further estimate
the normalized activity scores for individual uncertain inputs ($\tilde{\nu}_{i,r}$; $r$=1 since the
active subspace is 1-dimensional)
using the components of the dominant eigenvector as shown in Algorithm~\ref{alg:grad}.
The activity scores for the 19 uncertain pre-exponents ($A_i$'s), estimated
using the perturbation-based and regression-based approaches are plotted in
Figure~\ref{fig:comp_as} (left).  Additionally, for the purpose of
verification, we compare these estimates with those obtained using the
normalized DGSMs in~\eqref{eq:ndgsm}, as reported in~\cite{Vohra:2018}, in
Figure~\ref{fig:comp_as} (right). 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_as}
  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
\caption{Left: A bar-graph of normalized activity scores ($\tilde{\nu}_{i,r}$'s) 
for the 19 uncertain pre-exponents ($A_i$'s); $r$
denotes the number of eigenvectors in the active subspace.
Right: A plot of the screening metric ($\tilde{\nu}_i$) for individual $A_i$'s 
(adapted from~\cite{Vohra:2018}).}
\label{fig:comp_as}
\end{center}
\end{figure}
%
The activity scores based on the two approaches for gradient estimation agree favorably 
with each other as well as those
based on the screening metric involving the DGSMs from~\cite{Vohra:2018}. It is observed that the uncertainty
associated with the ignition delay is largely due to the uncertainty in $A_9$ and $A_1$. Sensitivity towards
$A_{15}$ and $A_{17}$ is also found to be significant.

>>>>

As mentioned earlier, the model output $f(\bm{\xi})$ i.e. the ignition delay in the
H$_2$/O$_2$ reaction in this case, varies
 predominantly in the active subspace. Hence, 
$f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
the subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
regarded as the \textit{sufficient summary plot} (SSP), obtained using the perturbation-based and
regression-based gradient estimates are
compared in Figure~\ref{fig:comp_ssp}.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp}
\caption{An illustrative comparison of the SSPs generated using the 
perturbation-based and the regression-based strategies for computing the active subspace.}
\label{fig:comp_ssp}
\end{center}
\end{figure}
%
The two SSPs are observed to be in excellent agreement with each other. Moreover, it is interesting to note
that the response in ignition
delay based on the considered probability distributions for $A_i$'s is observed to be approximately linear in the
active subspace. This observation further indicates that the active subspace is 1-dimensional in this case.

We further estimate
the normalized activity scores for individual uncertain inputs ($\tilde{\nu}_{i,r}$; $r$=1 since the
active subspace is 1-dimensional)
using the components of the dominant eigenvector as shown in Algorithm~\ref{alg:grad}.
The activity scores for the 19 uncertain pre-exponents ($A_i$'s), estimated
using the perturbation-based and regression-based approaches are plotted in
Figure~\ref{fig:comp_as}.  
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_as}
%  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
\caption{Left: A bar-graph of normalized activity scores ($\tilde{\nu}_{i,r}$'s) 
for the 19 uncertain pre-exponents ($A_i$'s); $r$
denotes the number of eigenvectors in the active subspace.}
\label{fig:comp_as}
\end{center}
\end{figure}
%
The activity scores based on the two approaches for gradient estimation agree favorably 
with each other as well as those
based on the screening metric involving the DGSMs in~\cite{Vohra:2018}. It is observed that the uncertainty
associated with the ignition delay is largely due to the uncertainty in $A_9$ and $A_1$. Sensitivity towards
$A_{15}$ and $A_{17}$ is also found to be significant.
